---
title: "Fuel-Substitution Cost Effectiveness Analysis, HVAC and DHWHeat Pump Measures"
output:
  word_document: default
  html_document: default
date: "2024-04-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(compare)
library(rmarkdown)
library (dplyr)
library (scales)
library (kableExtra)

```
## Background 

### The Problem:
Energy Efficiency (EE) uses the Total Resource Cost (TRC) test to determine what EE resources to procure. TRC represents costs and benefits of EE projects to the Utility and Participant. Unfortunately, most Fuel-Substitution measures have very low TRC values under EE's current Cost-Effectiveness framework which means Program Administrators struggle to implement them. 

ED recognizes that the TRC currently does not include the participant Utility Bill Savings Benefits or the Costs of Infrastructure upgrades often involved in Fuel-Substitution projects. 

### Research Question:
If EE factored Participant Utility Bill Savings Benefits or the Costs of Infrastructure upgrades, how would this impact the TRC of Fuel-Sub projects?

To explore this question, Calculate Gas Benefits with Utility Bill Savings, and Total Costs with Infrastructure Costs of the Fuel-Sub Measures claimed in PG&E's 2023 Plug Load and Appliance, and Upstream HVAC Programs in Single Family Homes.

### Inputs

```{r, echo=TRUE}
#Selections
MeasAppType <- 'NR'
BldgType <- 'SFm'
SZ <- 'med'
ClimateZone <- 12
HeatPumpHVAC_res <- 'SWHC045'  
HPWH_res <- 'SWWH025'
EU_HP_HVAC<-'DEER:HVAC_Eff_HP'
EU_HPWH<-'DEER:Res_ClothesDishWasher'
EUL_Claims_HeatPumpHVAC_res <- 15 #2024 Study recommends 23 years
EUL_Claims_HPWH_res <- 10 #2024 Study recommends 20 years
EUL_New_HeatPumpHVAC_res <- 23 #2024 Study recommends 23 years
EUL_New_HPWH_res <- 20 #2024 Study recommends 20 years
#Constants
PGE_WACC_A <- 0.0734 #Source:ACC, discount Rate
PGE_WACC_Q <- PGE_WACC_A/4
Divider_discount <- PGE_WACC_Q + 1
```

## Benefits Analysis
### Data Sources

The data inputs are shown below:

• Claims data was pulled from CEDARS in April 2024: https://cedars.sound-data.com/reports/download-record-level-report/claims/2023/

• Load Shapes profiles with Avoided Costs are from CEDARS: https://file.ac/l1-GqhWF8OU/

• Utility Bill Impacts were developed by Guidehouse

The data from the online sources were truncated and renamed to simplify this analysis; the names of these files are listed below. 

```{r, echo=TRUE}

#Read in Fuel Sub Claims 2023 Q1-Q4
FS23 <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\23_FS_Claims_PGE.xlsx") #Normal Replacement Only (No New Construction)

#Read in Load Shapes (accessible at https://file.ac/l1-GqhWF8OU/)
Gas_LS_QRT <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\G_LS_QRT.xlsx") #Quarterly LS

#Read in Bill Impacts Data
Utility_Bill_Ben_CARE <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\data_TH\\res_bill_impacts_PGE_Monthly_CARE.xlsx")
Utility_Bill_Ben_Market <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\data_TH\\res_bill_impacts_PGE_Monthly_Market.xlsx")
```


```{r, echo=FALSE}
#Read in Claim Yr Quarters
Quarters <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\Qrt.xlsx")
FS23 <- as.data.frame(FS23)

#Total Energy: by Measure & CZ
df_HeatPumpHVAC_res<- filter(FS23, CZ == ClimateZone, MeasureID == HeatPumpHVAC_res,`Measure Application Type` == MeasAppType, `Building Type` == BldgType )
df_HeatPumpHVAC_res  <- df_HeatPumpHVAC_res %>% mutate(Therm_TotalAnnual=df_HeatPumpHVAC_res$`Unit Therm First Baseline` * df_HeatPumpHVAC_res$`Number of Units`)
HeatPumpHVAC_res_therm<- c(sum(df_HeatPumpHVAC_res$`Therm_TotalAnnual`))

df_HPWH_res<- filter(FS23, CZ == ClimateZone, MeasureID == HPWH_res, `Measure Application Type` == MeasAppType, `Building Type` == BldgType)
df_HPWH_res<-df_HPWH_res%>% mutate(Therm_TotalAnnual=df_HPWH_res$`Unit Therm First Baseline` * df_HPWH_res$`Number of Units`)
HPWH_res_therm<- c(sum(df_HPWH_res$`Therm_TotalAnnual`))

#Add Bill Savings Bens
#filter, Size, CZ
#x3 to get quarterly
Num_Measures <- FS23 %>% count(MeasureID)
Num_HPWH <- filter(Num_Measures, MeasureID == HPWH_res ) 
Num_HVAC <- filter(Num_Measures, MeasureID == HeatPumpHVAC_res ) 

Utility_Bill_M <- filter(Utility_Bill_Ben_Market, `CZ` == ClimateZone, `Building Type` == BldgType, `Size` == SZ)

Utility_Bill_HVAC_M <- (Utility_Bill_M$HP_HVAC*12*EUL_New_HeatPumpHVAC_res)/(1+PGE_WACC_A)^EUL_New_HeatPumpHVAC_res
Utility_Bill_HVAC_M_Total <- (Num_HVAC$n*Utility_Bill_HVAC_M)

Utility_Bill_HPWH_M <- (Utility_Bill_M$HPWH*12*EUL_New_HPWH_res)/(1+PGE_WACC_A)^EUL_New_HPWH_res
Utility_Bill_HPWH_M_Total <- (Num_HPWH$n*Utility_Bill_HPWH_M)


Utility_Bill_C <- filter(Utility_Bill_Ben_CARE, `CZ` == ClimateZone, `Building Type` == BldgType, `Size` == SZ)

Utility_Bill_HVAC_C <- (Utility_Bill_C$HP_HVAC*12*EUL_New_HeatPumpHVAC_res)/(1+PGE_WACC_A)^EUL_New_HeatPumpHVAC_res
Utility_Bill_HVAC_C_Total <- (Num_HVAC$n*Utility_Bill_HVAC_C)

Utility_Bill_HPWH_C <- (Utility_Bill_C$HPWH*12*EUL_New_HPWH_res)/(1+PGE_WACC_A)^EUL_New_HPWH_res
Utility_Bill_HPWH_C_Total <- (Num_HPWH$n*Utility_Bill_HPWH_C)

Claims_GasBen_HeatPumpHVAC_res<- c(sum(df_HeatPumpHVAC_res$`Gas Benefits`))
Claims_GasBen_HPWH_res<- c(sum(df_HPWH_res$`Gas Benefits`))
```

### Calculate Benefits: Gas Commodity, T&D, GHG, Utility Bill
```{r, echo= TRUE}
#Calculated Quarterly Avoided Costs (energy x LS) over EUL of the Claims
#Generate the Claim Year Quarters
EUL_Claims_HeatPumpHVAC_res <-  list(Quarters$EUL_15) 
EUL_Claims_HeatPumpHVAC_res<- as.data.frame(EUL_Claims_HeatPumpHVAC_res)

EUL_Claims_HPWH_res <-  list(Quarters$EUL_10) 
colnames(EUL_Claims_HeatPumpHVAC_res) <- c('CYQtr')
EUL_Claims_HPWH_res<- as.data.frame(EUL_Claims_HPWH_res)
colnames(EUL_Claims_HPWH_res) <- c('CYQtr')

EUL_New_HeatPumpHVAC_res <-  list(Quarters$EUL_23) 
EUL_New_HeatPumpHVAC_res<- as.data.frame(EUL_New_HeatPumpHVAC_res)

EUL_New_HPWH_res <-  list(Quarters$EUL_20) 
colnames(EUL_New_HeatPumpHVAC_res) <- c('CYQtr')
EUL_New_HPWH_res<- as.data.frame(EUL_New_HPWH_res)
colnames(EUL_New_HPWH_res) <- c('CYQtr')

#Total Therm Savings from Claims Data 
Therm_TotalAnnual_HVAC <- c(sum(df_HeatPumpHVAC_res$`Therm_TotalAnnual`))
Therm_TotalAnnual_HPWH <- c(sum(df_HPWH_res$`Therm_TotalAnnual`))

#Generate the NPV of Commodity and T&D Avoided Costs
generate_qtrly_npv_1 <- function(load_shape, 
                               considered_qtrs, discount_rate, Therms) {
  filtered_load_shape <- filter(load_shape, 
                                Qtr %in% considered_qtrs$CYQt) %>% 
    mutate(Qtr_step = 1:n()) %>% 
    mutate(Qtr_ID = Qtr_step-1) %>% 
    mutate(Total_Discounted = (Total/(discount_rate^Qtr_ID)))
  
  result_1 <- c(sum(filtered_load_shape$'Total_Discounted'*Therms))
  return (result_1)
}

GasComTD_HeatPumpHVAC_res <- 
  generate_qtrly_npv_1(Gas_LS_QRT, 
                     EUL_Claims_HeatPumpHVAC_res, Divider_discount,
                     Therm_TotalAnnual_HVAC)

GasComTD_HPWH_res <- 
  generate_qtrly_npv_1(Gas_LS_QRT, EUL_Claims_HPWH_res,
                     Divider_discount, Therm_TotalAnnual_HPWH )

GasComTD_HeatPumpHVAC_res_new <- 
  generate_qtrly_npv_1(Gas_LS_QRT, 
                     EUL_New_HeatPumpHVAC_res, Divider_discount,
                     Therm_TotalAnnual_HVAC)

GasComTD_HPWH_res_new <- 
  generate_qtrly_npv_1(Gas_LS_QRT, EUL_New_HPWH_res,
                     Divider_discount, Therm_TotalAnnual_HPWH )


```

```{r, echo=TRUE}
 #Calculate kWhe of the Therm savings so we can use the Electric GHG Avoided Costs (the Gas GHG Avoided Costs were not generated in the latest data set)
 Gas_kwhe_HPWH <- Therm_TotalAnnual_HPWH*29.3
 Gas_kwhe_HVAC <- Therm_TotalAnnual_HVAC*29.3
 kWh_LS_QRT <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\E_LS_QRT_22.xlsx") #Quarterly LS
 
 #Generate the NPV of CO2 Avoided Costs
 generate_qtrly_npv_2 <- function(load_shape, climate_zone, end_use,
                               considered_qtrs, discount_rate, gas_kwhe) {
   filtered_load_shape <- filter(load_shape, CZ == climate_zone, EU == end_use,
                                 Qtr %in% considered_qtrs$CYQt) %>%
     mutate(Qtr_step = 1:n()) %>%
     mutate(Qtr_ID = Qtr_step-1) %>%
     mutate(CO2_Discounted = (CO2/(discount_rate^Qtr_ID)))
 
   result <- 2*c(sum(filtered_load_shape$'CO2_Discounted'*gas_kwhe))
  return (result)
}

#Gas GHG Adder
# Gas_GHG_Adder <- 114
# 
# Gas_Ben_GHG_HPWH <- Therm_TotalAnnual_HPWH * Gas_GHG_Adder
# percent_inc_GasBen_HPWH <- (Gas_Ben_GHG_HPWH - Calc_GasBen_HPWH_res)/ Calc_GasBen_HPWH_res
# 
# Gas_Ben_GHG_HPHVAC <- Therm_TotalAnnual_HVAC * Gas_GHG_Adder
# percent_inc_GasBen_HPHVAC <- (Gas_Ben_GHG_HPHVAC - Calc_GasBen_HeatPumpHVAC_res)/ Calc_GasBen_HeatPumpHVAC_res
# 
# 
 GasCO2_HeatPumpHVAC_res <-   generate_qtrly_npv_2 (kWh_LS_QRT, ClimateZone, EU_HP_HVAC,EUL_Claims_HeatPumpHVAC_res,Divider_discount, Gas_kwhe_HVAC )
 GasCO2_HPWH_res <-   generate_qtrly_npv_2 (kWh_LS_QRT, ClimateZone, EU_HPWH,EUL_Claims_HPWH_res,Divider_discount, Gas_kwhe_HPWH)
   
GasCO2_HeatPumpHVAC_res_new <-  generate_qtrly_npv_2 (kWh_LS_QRT, ClimateZone, EU_HP_HVAC,EUL_New_HeatPumpHVAC_res,Divider_discount, Gas_kwhe_HVAC)
 GasCO2_HPWH_res_new <-  generate_qtrly_npv_2(kWh_LS_QRT, ClimateZone, EU_HPWH,EUL_New_HPWH_res,Divider_discount, Gas_kwhe_HPWH)

#Calculate Total Gas Benefits
Calc_GasBen_HeatPumpHVAC_res <- GasComTD_HeatPumpHVAC_res # GasCO2_HeatPumpHVAC_res+ 
Calc_GasBen_HPWH_res <-  GasComTD_HPWH_res #GasCO2_HPWH_res +

Calc_GasBen_HeatPumpHVAC_res_new <-  GasCO2_HeatPumpHVAC_res_new+ GasComTD_HeatPumpHVAC_res_new
Calc_GasBen_HPWH_res_new <- GasCO2_HPWH_res_new + GasComTD_HPWH_res_new

```


```{r, echo=FALSE}
HP_HVAC <- dollar(Calc_GasBen_HeatPumpHVAC_res)
HP_HVAC_Bill <-dollar(Utility_Bill_HVAC_M_Total)
BillImpact_HVAC<- percent((Utility_Bill_HVAC_M_Total)/Calc_GasBen_HeatPumpHVAC_res_new)

HPWH <- dollar(Calc_GasBen_HPWH_res)
HPWH_Bill <-dollar(Utility_Bill_HPWH_M_Total)
BillImpact_HPWH<- percent((Utility_Bill_HPWH_M_Total)/Calc_GasBen_HPWH_res_new)

```

The Lifecycle FS NR Heat Pump HVAC Gas Benefits of the 2023 PGE claims are **`r HP_HVAC`**, and the total Lifecycle, Market-Rate Utility Bill savings for all mid-sized single family homes in these same claims are **`r HP_HVAC_Bill`**. 

The Lifecycle FS NR HPWH Gas Benefits of the 2023 PGE claims are **`r HPWH`**, and the total Lifecycle, Market-Rate Utility Bill savings for all mid-sized single family homes in these same claims are **`r HPWH_Bill`**. 


```{r, echo=FALSE}
# Table
Results_gas <- data.frame(
  Measure = c("HVAC Heat Pump", "Heat Pump Water Heater"),
  "Gas Benefits" = c(HP_HVAC, HPWH),
  "Bill Market Savings" = c(HP_HVAC_Bill,HPWH_Bill)
  
)
```

```{r, echo=FALSE}

knitr::kable(Results_gas, caption = "Total Benefits for Fuel Sub Measure Claims", 
             col.names = c("Measure", "Gas Benefits","Bill Market Savings"))

```





## Cost Analysis


### Data Sources

The data inputs are shown below:

• Claims data was pulled from CEDARS in April 2024: https://cedars.sound-data.com/reports/download-record-level-report/claims/2023/

• Load Shapes profiles with Avoided Costs are from CEDARS: https://file.ac/l1-GqhWF8OU/

• Infrastructure Costs were based on ED Study (LINK TO PDA)

The data from the online sources were truncated and renamed to simplify this analysis; the names of these files are listed below. 

```{r, echo=TRUE}
#Read in Fuel Sub Claims 2023 Q1-Q4
FS23 <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\23_FS_Claims_PGE.xlsx") #Normal Replacement Only (No New Construction)

#Read in Load Shapes (accessible at https://file.ac/l1-GqhWF8OU/)
kWh_LS_QRT <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\E_LS_QRT_22.xlsx") #Quarterly LS

#Read in Infrastructure Costs
Infra_Costs <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\data_TH\\FS_InfraCosts.xlsx") #input as full cost to install Qtr. No Discounting Needed
```

```{r, echo=FALSE}
FS23 <- as.data.frame(FS23)
#Read in Claim Yr Quarters
Quarters <- read_excel("C:\\Users\\mmh\\R  Programming\\subtests\\FS_NewBens\\Qrt.xlsx")
#Total Energy: by Measure & CZ
df_HeatPumpHVAC_res<- filter(FS23, CZ == ClimateZone, MeasureID == HeatPumpHVAC_res, `Measure Application Type` == MeasAppType, `Building Type` == BldgType) 
df_HeatPumpHVAC_res  <- df_HeatPumpHVAC_res %>% mutate(kWh_TotalAnnual=df_HeatPumpHVAC_res$`Unit kWh First Baseline` * df_HeatPumpHVAC_res$`Number of Units`)
HeatPumpHVAC_res_kWh<- c(sum(df_HeatPumpHVAC_res$`kWh_TotalAnnual`))

df_HPWH_res<- filter(FS23, CZ == ClimateZone, MeasureID == HPWH_res,`Measure Application Type` == MeasAppType, `Building Type` == BldgType)
df_HPWH_res<-df_HPWH_res %>% mutate(kWh_TotalAnnual=df_HPWH_res$`Unit kWh First Baseline` * df_HPWH_res$`Number of Units`)
HPWH_res_kWh<- c(sum(df_HPWH_res$`kWh_TotalAnnual`))

#Claimed Electric Costs and Gas Benefits
#df_HeatPumpHVAC_res <- lapply(df_HeatPumpHVAC_res,as.numeric)
#df_HPWH_res <- lapply(df_HPWH_res,as.numeric)


Claims_ElecCost_HeatPumpHVAC_res<- c(sum(df_HeatPumpHVAC_res$`Electric Supply Cost Gross`))
Claims_ElecCost_HPWH_res<- c(sum(df_HPWH_res$`Electric Supply Cost Gross`))

#Calculated Quarterly Electric ACs (energy x LS) over EUL_Claims
#Filter for CZ, End Use, CLaimYearQrt
EUL_Claims_HeatPumpHVAC_res <-  list(Quarters$EUL_15) 
EUL_Claims_HeatPumpHVAC_res<- as.data.frame(EUL_Claims_HeatPumpHVAC_res)
colnames(EUL_Claims_HeatPumpHVAC_res) <- c('CYQtr')

EUL_Claims_HPWH_res <-  list(Quarters$EUL_10) 
EUL_Claims_HPWH_res<- as.data.frame(EUL_Claims_HPWH_res)
colnames(EUL_Claims_HPWH_res) <- c('CYQtr')


#Calculated Quarterly Electric ACs (energy x LS) over EUL_Claims
#Filter for CZ, End Use, CLaimYearQrt
EUL_New_HeatPumpHVAC_res <-  list(Quarters$EUL_23) 
EUL_New_HeatPumpHVAC_res<- as.data.frame(EUL_New_HeatPumpHVAC_res)
colnames(EUL_New_HeatPumpHVAC_res) <- c('CYQtr')

EUL_New_HPWH_res <-  list(Quarters$EUL_20) 
EUL_New_HPWH_res<- as.data.frame(EUL_New_HPWH_res)
colnames(EUL_New_HPWH_res) <- c('CYQtr')



```


### Calculate Costs: Electric Supply Costs & Infrastructure Costs

```{r, echo= TRUE}

kWh_TotalAnnual_HVAC <- c(sum(df_HeatPumpHVAC_res$`kWh_TotalAnnual`))
kWh_TotalAnnual_HPWH <- c(sum(df_HPWH_res$`kWh_TotalAnnual`))

#Generate the NPV of CO2 Avoided Costs
generate_qtrly_npv_e <- function(load_shape, climate_zone, end_use, 
                                 considered_qtrs, discount_rate,
                                 kWh_TotalAnnual, AC) {
  filtered_load_shape <- filter(load_shape, CZ == climate_zone, EU == end_use, 
                                Qtr %in% considered_qtrs$CYQt) %>% 
    mutate(Qtr_step = 1:n()) %>% 
    mutate(Qtr_ID = Qtr_step-1) %>% 
    mutate(AC_type_discounted = (!!sym(AC)/(discount_rate^Qtr_ID)))
  
 result3 <- c(sum(filtered_load_shape$AC_type_discounted*kWh_TotalAnnual))
  return (result3)
 
}
#old EUL
AC_Sum_Gen_E_HeatPumpHVAC_res <-
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HP_HVAC,
                     EUL_Claims_HeatPumpHVAC_res, Divider_discount,
                     kWh_TotalAnnual_HVAC,'Gen')
AC_Sum_TD_E_HeatPumpHVAC_res <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HP_HVAC, 
                    EUL_Claims_HeatPumpHVAC_res, Divider_discount,
                    kWh_TotalAnnual_HVAC,'TD')
AC_Sum_Co2_E_HeatPumpHVAC_res <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HP_HVAC, 
                     EUL_Claims_HeatPumpHVAC_res, Divider_discount,
                     kWh_TotalAnnual_HVAC,'CO2')


AC_Sum_Gen_E_HPWH_res <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HPWH, EUL_Claims_HPWH_res,
                     Divider_discount, kWh_TotalAnnual_HPWH,'Gen')
AC_Sum_TD_E_HPWH_res <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HPWH, EUL_Claims_HPWH_res,
                     Divider_discount, kWh_TotalAnnual_HPWH,'TD')
AC_Sum_Co2_E_HPWH_res <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HPWH, EUL_Claims_HPWH_res,
                     Divider_discount, kWh_TotalAnnual_HPWH,'CO2')

#new EUL
AC_Sum_Gen_E_HeatPumpHVAC_res_new <-
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HP_HVAC,
                     EUL_New_HeatPumpHVAC_res, Divider_discount,
                     kWh_TotalAnnual_HVAC,'Gen')
AC_Sum_TD_E_HeatPumpHVAC_res_new <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HP_HVAC, 
                    EUL_New_HeatPumpHVAC_res , Divider_discount,
                    kWh_TotalAnnual_HVAC,'TD')
AC_Sum_Co2_E_HeatPumpHVAC_res_new <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HP_HVAC, 
                     EUL_New_HeatPumpHVAC_res , Divider_discount,
                     kWh_TotalAnnual_HVAC,'CO2')


AC_Sum_Gen_E_HPWH_res_new <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HPWH, EUL_New_HPWH_res,
                     Divider_discount, kWh_TotalAnnual_HPWH,'Gen')
AC_Sum_TD_E_HPWH_res_new <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HPWH, EUL_New_HPWH_res,
                     Divider_discount, kWh_TotalAnnual_HPWH,'TD')
AC_Sum_Co2_E_HPWH_res_new <- 
  generate_qtrly_npv_e(kWh_LS_QRT, ClimateZone, EU_HPWH, EUL_New_HPWH_res,
                     Divider_discount, kWh_TotalAnnual_HPWH,'CO2')

#Calculate Total Electric Costs 
Calc_ElecCost_HeatPumpHVAC_res <-  (AC_Sum_Gen_E_HeatPumpHVAC_res + AC_Sum_TD_E_HeatPumpHVAC_res + AC_Sum_Co2_E_HeatPumpHVAC_res)*(-1)
Calc_ElecCost_HPWH_res <- (AC_Sum_Gen_E_HPWH_res + AC_Sum_TD_E_HPWH_res + AC_Sum_Co2_E_HPWH_res)*(-1)


Calc_ElecCost_HeatPumpHVAC_res_new <-  (AC_Sum_Gen_E_HeatPumpHVAC_res_new  + AC_Sum_TD_E_HeatPumpHVAC_res_new  + AC_Sum_Co2_E_HeatPumpHVAC_res_new  )*(-1)
Calc_ElecCost_HPWH_res_new  <- (AC_Sum_Gen_E_HPWH_res_new  + AC_Sum_TD_E_HPWH_res_new  + AC_Sum_Co2_E_HPWH_res_new )*(-1)
```
``` {r echo=TRUE}

#Add Infra Costs
Num_Measures <- FS23 %>% count(MeasureID)
Num_HPWH <- filter(Num_Measures, MeasureID == HPWH_res ) 
Num_HVAC <- filter(Num_Measures, MeasureID == HeatPumpHVAC_res ) 
InfraCosts <- filter(Infra_Costs, Sector == 'res')

HP_HVAC_InfraCosts <- InfraCosts$HP_HVAC*Num_HVAC$n
HPWH_InfraCosts <- InfraCosts$HPWH*Num_HPWH$n

HP_HVAC_InfC <- dollar(HP_HVAC_InfraCosts)
HPWH_InfC <- dollar(HPWH_InfraCosts)
TRC_Cost_HPHVAC <- dollar (c(sum(df_HeatPumpHVAC_res$'TRC Cost Gross')))
TRC_Cost_HPWH<- dollar(c(sum(df_HPWH_res$'TRC Cost Gross')))
```
The Lifecycle FS NR Heat Pump HVAC Total Costs for the 2023 PGE claims are **`r TRC_Cost_HPHVAC`**, and the residential infrastructure costs are  **`r HP_HVAC_InfC `**. 

The Lifecycle FS NR HPWH Gas Benefits of the 2023 PGE claims are **`r TRC_Cost_HPWH`**, and the and the residential infrastructure costs are **`r HPWH_InfC`**. 

```{r, echo=FALSE}
# Table
Results_costs <- data.frame(
  Measure = c("HVAC Heat Pump", "Heat Pump Water Heater"),
  "TRC Costs" = c(TRC_Cost_HPHVAC,TRC_Cost_HPWH),
  "Infrastructure Costs" = c(HP_HVAC_InfC,HPWH_InfC)
)
```

```{r, echo=FALSE}

knitr::kable(Results_costs, caption = "Total Costs for Fuel Sub Measure Claims",
             col.names = c("Measure", "TRC Costs","Infrastructure Costs"))

```



## Results: 

``` {r echo=TRUE}
#Clacualte increases to TRC Benefits

Claims_HP_HVAC <- filter(FS23, CZ == ClimateZone, MeasureID == HeatPumpHVAC_res, `Measure Application Type` == MeasAppType, `Building Type` == BldgType)
TRC_Claims_HP_HVAC <- mean(Claims_HP_HVAC$TRC)

Claims_HPWH <- filter(FS23, CZ == ClimateZone, MeasureID == HPWH_res, `Measure Application Type` == MeasAppType, `Building Type` == BldgType)
TRC_Claims_HPWH <- mean(Claims_HPWH$TRC)

TRC_New_HP_HVAC_EUL <- TRC_Claims_HP_HVAC * (1+((Calc_GasBen_HeatPumpHVAC_res_new-Calc_GasBen_HeatPumpHVAC_res)/Calc_GasBen_HeatPumpHVAC_res))
TRC_New_HPWH_EUL <- TRC_Claims_HPWH * (1+ ((Calc_GasBen_HPWH_res_new-Calc_GasBen_HPWH_res)/Calc_GasBen_HPWH_res))

#factor in elec costs increase to TRC

multiplier_elec_cost <- c(sum(df_HeatPumpHVAC_res$'Electric Supply Cost')) / c(sum(df_HeatPumpHVAC_res$'TRC Cost Gross'))
percent_inc_Costs_HPHVAC_EUL <-((Calc_ElecCost_HeatPumpHVAC_res_new-Calc_ElecCost_HeatPumpHVAC_res)/Calc_ElecCost_HeatPumpHVAC_res)* multiplier_elec_cost
TRC_EUL_Net_HPHVAC <-TRC_New_HP_HVAC_EUL - (TRC_New_HP_HVAC_EUL*percent_inc_Costs_HPHVAC_EUL)

multiplier_elec_cost_2 <- c(sum(df_HPWH_res$'Electric Supply Cost')) / c(sum(df_HPWH_res$'TRC Cost Gross'))
percent_inc_Costs_HPWH_EUL <- ((Calc_ElecCost_HPWH_res_new-Calc_ElecCost_HPWH_res)/Calc_ElecCost_HPWH_res)*multiplier_elec_cost_2
TRC_EUL_Net_HPWH<- TRC_New_HPWH_EUL - (TRC_New_HPWH_EUL*percent_inc_Costs_HPWH_EUL)

#Bill Impacts

Bill_EUL_HVAC <- Utility_Bill_HVAC_M_Total+Calc_GasBen_HeatPumpHVAC_res_new
Bill_EUL_HPWH <-Utility_Bill_HPWH_M_Total+ Calc_GasBen_HPWH_res_new
TRC_HP_HVAC_EUL_Bills <- TRC_EUL_Net_HPHVAC * (1+((Bill_EUL_HVAC-Calc_GasBen_HeatPumpHVAC_res_new)/Calc_GasBen_HeatPumpHVAC_res_new))
TRC_HPWH_EUL_Bills <- TRC_EUL_Net_HPWH  * (1+ ((Bill_EUL_HPWH-Calc_GasBen_HPWH_res_new)/Calc_GasBen_HPWH_res_new))


```



```{r, echo=FALSE}

#Decrease in TRC from above (i.e., due to bill savings and gas bens inc) due to  Infra Costs


#what % is InfraCost of the Total Costs with EUL inc  .... compare to percent_inc_Costs_xxxxx_EUL
percent_inc_Costs_HPHVAC_Infra <-HP_HVAC_InfraCosts/ (c(sum(df_HeatPumpHVAC_res$'Electric Supply Cost'))+c(sum(df_HeatPumpHVAC_res$'TRC Cost Gross')))
TRC_Bill_EUL_Net_Infra_HPHVAC <- TRC_HP_HVAC_EUL_Bills  - (TRC_HP_HVAC_EUL_Bills *percent_inc_Costs_HPHVAC_Infra)

percent_inc_Costs_HPWH_Infra <-HPWH_InfraCosts/ (c(sum(df_HPWH_res$'Electric Supply Cost'))+c(sum(df_HPWH_res$'TRC Cost Gross')))
TRC_Bill_EUL_Net_Infra_HPWH <- TRC_HPWH_EUL_Bills - (TRC_HPWH_EUL_Bills*percent_inc_Costs_HPWH_Infra )


#Table
TRC_IMPACT <- data.frame(
  Measure = c("HVAC Heat Pump", "Heat Pump Water Heater"),
  TRC_average = c(TRC_Claims_HP_HVAC,TRC_Claims_HPWH),
  TRC_inc_eul = c(TRC_EUL_Net_HPHVAC,TRC_EUL_Net_HPWH),
  TRC_inc_bill = c(TRC_HP_HVAC_EUL_Bills,TRC_HPWH_EUL_Bills),
  TRC_dec_infracosts = c(TRC_Bill_EUL_Net_Infra_HPHVAC,TRC_Bill_EUL_Net_Infra_HPWH)
)
TRC_IMPACT$TRC_average<- sprintf("%.2f", TRC_IMPACT$TRC_average)
TRC_IMPACT$TRC_inc_eul<- sprintf("%.2f", TRC_IMPACT$TRC_inc_eul)
TRC_IMPACT$TRC_inc_bill<- sprintf("%.2f", TRC_IMPACT$TRC_inc_bill)
TRC_IMPACT$TRC_dec_infracosts<- sprintf("%.2f", TRC_IMPACT$TRC_dec_infracosts)
knitr::kable(TRC_IMPACT, caption = "Average Total Resource Costs (TRC) for Fuel-Sub Measures",
             col.names = c("Measure", "TRC Claims", "TRC, Increased from EUL Study","TRC, Increased from EUL Study & Bill Impacts", "TRC, Increased from EUL Study & Bill Impacts, Decreased from Infra Cost"))
```



